### 2018-01-09 Manuscript/LogSpaceDistributionsCytometryA.tex
Small changes

\begin{figure}[!ht]
 \centering
%  \includegraphics[clip,width=1\hsize]{Fig1.tif}
  \caption{\textbf{Log-normal mixture showing the mismatch in the number of peaks.} \textbf{Top row, red:} $P(\log I)$ vs. $\log I$ normalized to max. \textbf{Middle row, magenta:} $Q(I)$ vs. $I$ normalized to max, and plotted on a narrower range. \textbf{Bottom row, blue:} $Q(I)$ vs. $\log I$ normalized to max, rescaling the x-axis as in Ref.~\cite{Novo2008}. In the central column where $\sigma_1=\sigma_2=0.8$, $P(\log I)$ shows explicit bimodality whereas $Q(I)$ is unimodal. \textbf{Right column:} in this case, the variance $\sigma^2$ is large enough (see Eq.~\ref{eq:Plogy3Maxima}) that $P(\log I)$ has only one mode even though it is modeled as a mixture. In all cases: $I_1=100,\,I_2=1000,\,\alpha=0.5$ varying $\sigma_{1,2}=\{0.4,0.8,1.2\}$ (left to right columns, respectively). Leftmost column shows axes. }
  \label{fig:PlotsLinearLog}
\end{figure}

@@

\smallskip
$A(y)$ is the ratio of the two Gaussians in Eq.~\ref{eq:LogNormalMix} and is therefore always non-negative; this implies that any extremum $y_*$ must satisfy $B_1(y_{*q})\ge 0$ and $B_3(y_{*p})\ge 0$. We note that the condition for extrema in $Q(I)$ requires that $\frac{d}{dy}P(y)=P(y)$ whereas, of course, extremizing $P(\log I)$ sets its derivative to zero, demonstrating the fact that the loci of the modes for $P(\log I)$ and $Q(I)$ are manifestly different. The region where the log-space distribution shows a second mode occurs when for $B_3(y_{*p})=A(y_{*p})$ admits three solutions whereas $B_1(y_{*q})=A(y_{*q})$ admits only one solution. Given that these equations are transcendental, a graphical way to asses the number of solutions is to plot $A(y),B_1(y),B_3(y)$ and count the number of times $B_1$ and $B_3$ intersect $A$.\smallskip

@@

\smallskip
\begin{itemize}
\item All MATLAB scripts used to generate the plots and analyze the data in 
github.com/AmirErez/BimodalLogspaceCytA/tree/master/Scripts 
\item Al data used in the manuscript and accompanying MIFlowCyt checklists come from two experiments:
	\begin{enumerate}
	\item 20140920-OT1-dynamics - used for Fig. 3 in the manuscript.
	\item FeinermanScience2008 - previously published in Feinerman et al. Science Vol. 321, Issue 5892, pp. 1081-1084 (2008). 
	\end{enumerate}
These data with their respective MIFlowCyt checklists are available in 
github.com/AmirErez/BimodalLogspaceCytA/tree/master/Data 
\end{itemize} 
All are accessible in 
https://github.com/AmirErez/BimodalLogspaceCytA
We briefly sketch the methods for each of the two experiments:
\subsubsection*{Experiment 1:} Primary mouse T lymphocytes (OT-1) were activated ex vivo and cultured. OT-1 cells were activated by peptide (SIINFEKL) treated and irradiated (3,000 RAD) primary antigen presenting cells (APCs) from a C57BL/6 mouse. OT-1 cells were labeled with an amine-reactive dye, CTV, according to the manufacture’s protocol (Molecular Probes) for in silico identification. Cells were allowed to rest for one hour after CTV staining, and then distributed in a 96-well v-bottom plate. Cells were then incubated with unique doses of SRC inhibitor Dasatinib at a temperature of 37C for 5 minutes. After which,  we added the peptide pulsed APCs (10 APCs to 1 OT-1 T cell) and pelleted mixture of cells by centrifugation for 10 s at 460 rcf at room temperature.  The pellet was then incubated at a temperature of 37C for 10 minutes, followed by 15 minutes of chemical fixing with 2\% PFA on ice and then permeabilization with ice cold 90\% MeOH. Samples were kept at a temperature of -20C until labeling for FCM.

\subsubsection*{Experiment 2:} C57BL/6N splenocytes were pulsed for 2hr with 100nM SIINFEKL peptide, then irradiated (3000RAD), washed once and used as stimulator/feeder cells. OT-1 lymphocytes were harvested from auxiliary, brachial and inguinal lymph nodes as well as spleen (splenocytes were treated with ACK lysis buffer to remove red blood cells), and mixed with SIINFEKL-pulsed C57BL/6N splenocytes in complete RPMI. After two days, cells were expanded by diluting 2 fold into medium containing 100 pM IL-2. After four days, the cells were again expanded by 2 fold dilution into medium with IL-2. After one more day of culture, cells were harvested and spun through a 1.09 density Ficoll- Paque Plus gradient (GE Healthcare) to remove dead cells. Live cells were recovered, washed twice in complete medium and resuspended at 1 million/ml in complete medium with 100pM IL-2. Cells were used for experiments between 6 and 8 days after primary stimulation. The ppERK response to SRC and MEK inhibition was measured using primary OT-1 T-lymphocytes activated with RMA-S APCs. RMA-S cells were suspended in culture with 1 nM SIINFEKL peptide for 2 h at 37C, 5\% CO\textsubscript{2}, and on a rotator to guarantee mixing. During this time we labelled OT-1 cells with an amine-reactive dye, CTV, according to the manufacture’s protocol (Molecular Probes). We rested the OT-1 cells one hour after CTV staining, and then distributed them in a 96-well v-bottom plate. Each well was given various doses of SRC inhibitor and MEK inhibitor and kept at 37C for 5 min. Following the 5 min exposure to the inhibitors, we added the peptide pulsed RMA-S (10 RMA-S to 1 OT-1 T cell) and pelleted by centrifugation for 10 s at 460 rcf at room temperature. This step guaranteed that both cell types, OT-1 and RMA-S, came into contact. The cells were allowed to activate for 10 min, followed by fixing on ice in 2\% PFA, and then permeabilized and stored in 90\% MeOH at  -20C. APCs were pulsed with serial dilutions of OVA or variant peptides for 2 hr at 37C, then washed with T cell medium at the time of harvest, and resuspended with anti-CD8α (53-6.72)-Fab-coated T cells in their conditioned media in a V-bottom 96-well plate (Corning). Fab-coating was performed 10 min before cell use by incubating T cells with 10 μg/ml of Fab fragment. T:APC cell contacts were synchronized using a quick centrifugal spin (10s at 400g). Plates containing T:APC conjugates were placed on a water bath at 37C and incubated for 5 min. Supernatants were then discarded, T:APC conjugates disrupted by vortexing, and cells resuspended in ice-cold 4\% paraformaldehyde for 15 min. Cells were then permeabilized with ice-cold 90\% methanol for 15 min on ice, and washed twice with FACS buffer.

@@

\smallskip
In experiment 1, the T-cells were inhibited by the SRC inhibitor Dasatinib. In Fig.~\ref{fig:Experiment}(A) we see how a commercially-available analysis software (FlowJo \cite{FlowJo}) plots the distribution of ppERK in such an experiment, which clearly shows a bimodal structure. Fig.~\ref{fig:Experiment}(B) plots those same data, formally $P(\log I) := \hat{f}_{\log I}^{(8751)}$, the probability density estimated from the data by taking the logarithm of the intensity of 8751 cells and binning it in bins of size 0.2. Our procedure for estimating $P(\log I)$ gives the two modes as in FlowJo (red dots) whereas $Q(I)$ (derived using Eq.~\ref{eq:P2Q}) has a single mode (blue dots). We fit $P(\log I)$ as a Gaussian mixture. This is followed in Fig.~\ref{fig:Experiment}(C) by the same extrema analysis as in Fig.~\ref{fig:ExtremaSolutions}, revealing that indeed $Q(I)$ has a single maximum.
\smallskip

@@

@@

\subsection*{Relation between ERK1 level and activation status}
\begin{figure}[!htb]
 \centering
%   \includegraphics[width=1\linewidth]{Fig7.tif}  
  \caption{{\bf Test for weak dependence of ERK1 and activation status.} Whereas independence implies that $P(\log ERK1) = P(\log ERK1 \vert Activation)$, in fact there is a weak dependence between them regardless of whether we employ a vertical (red, defined by threshold value $ppERK_*$) or a diagonal (grey) gate (the diagonal gate showing a weaker dependence); this is observed directly by noting that the different distributions in Fig.~\ref{S1_Fig} do not lie on top of each other. To quantify this difference, the inset shows the mutual information between $P(\log ERK1)$ and $P(\log ERK1 | Activation)$, with the circle pointing out the particular concentration of stimulus (out of all concentrations used in Ref.~\cite{Feinerman2008}) we chose to plot in this example. The chosen concentration has the highest mutual information, {\it i.e.}, the lowest ability to discern between the active and inactive states.}
  %\caption{\textbf{By dividing ppERK readings by ERK1, we can ameliorate the mismatch between the two representations.} \textbf{(A)} $P(\log I)$ and $Q(I)$ vs. $\log I$ (dots: data, lines: gaussian mixture fit) together with \textbf{(B)} their extrema analysis, showing that the second mode in log ppERK does not exist if the data is linearly binned. \textbf{(C)} The same treatment but for $\tilde{I} = I_{ppERK}/I_{ERK1}$, \textbf{(D)} shows that both $\tilde{I}$ and $\log \tilde{I}$ have two modes, thus normalizing ppERK levels by total ERK1 maintains the bimodal structure both in $P(\log \tilde{I})$ and in $Q(\tilde{I})$. }
  \label{S1_Fig}
\end{figure}

>>>>

\begin{figure}[!ht]
 \centering
%  \includegraphics[clip,width=1\hsize]{Fig1.tif}
  \caption{\textbf{Log-normal mixture showing the mismatch in the number of peaks.} \textbf{Top row, red:} $P(\log I)$ vs. $\log I$ normalized to maximum probability. \textbf{Middle row, magenta:} $Q(I)$ vs. $I$ normalized to maximum probability, and plotted on a narrower range. \textbf{Bottom row, blue:} $Q(I)$ vs. $\log I$ normalized to maximum probability, rescaling the x-axis as in Ref.~\cite{Novo2008}. In the central column where $\sigma_1=\sigma_2=0.8$, $P(\log I)$ shows explicit bimodality whereas $Q(I)$ is unimodal. \textbf{Right column:} in this case, the variance $\sigma^2$ is large enough (see Eq.~\ref{eq:Plogy3Maxima}) that $P(\log I)$ has only one mode even though it is modeled as a mixture. In all cases: $I_1=100,\,I_2=1000,\,\alpha=0.5$ varying $\sigma_{1,2}=\{0.4,0.8,1.2\}$ (left to right columns, respectively). Leftmost column shows axes. }
  \label{fig:PlotsLinearLog}
\end{figure}

@@

\smallskip
$A(y)$ is the ratio of the two Gaussians in Eq.~\ref{eq:LogNormalMix} and is therefore always non-negative; this implies that any extremum $y_*$ must satisfy $B_1(y_{*q})\ge 0$ and $B_3(y_{*p})\ge 0$. In the \emph{Supporting information}, we derive that the condition for extrema in $Q(I)$ requires that $\frac{d}{dy}P(y)=P(y)$ whereas, of course, extremizing $P(\log I)$ sets its derivative to zero, demonstrating the fact that the loci of the modes for $P(\log I)$ and $Q(I)$ are manifestly different. The region where the log-space distribution shows a second mode occurs when for $B_3(y_{*p})=A(y_{*p})$ admits three solutions whereas $B_1(y_{*q})=A(y_{*q})$ admits only one solution. Given that these equations are transcendental, a graphical way to asses the number of solutions is to plot $A(y),B_1(y),B_3(y)$ and count the number of times $B_1$ and $B_3$ intersect $A$.\smallskip

@@

\smallskip
\begin{itemize}
\item All MATLAB scripts used to generate the plots and analyze the data in 
github.com/AmirErez/BimodalLogspaceCytA/tree/master/Scripts 
\item All data used in the manuscript and accompanying MIFlowCyt checklists come from two experiments:
	\begin{enumerate}
	\item 20140920-OT1-dynamics - used for Fig. 3 in the manuscript.
	\item FeinermanScience2008 - previously published in Feinerman et al. Science Vol. 321, Issue 5892, pp. 1081-1084 (2008). 
	\end{enumerate}
These data with their respective MIFlowCyt checklists are available in 
github.com/AmirErez/BimodalLogspaceCytA/tree/master/Data 
\end{itemize} 
All are accessible in 
https://github.com/AmirErez/BimodalLogspaceCytA
We briefly sketch the methods for each of the two experiments:
\subsubsection*{Experiment 1:} Primary mouse T lymphocytes (OT-1) were activated ex vivo and cultured. OT-1 cells were activated by peptide (SIINFEKL) treated and irradiated (3,000 RAD) primary antigen presenting cells (APCs) from a C57BL/6 mouse. OT-1 cells were labeled with an amine-reactive dye, CTV, according to the manufacture’s protocol (Molecular Probes) for in silico identification. Cells were allowed to rest for one hour after CTV staining, and then distributed in a 96-well v-bottom plate. Cells were then incubated with unique doses of SRC inhibitor Dasatinib at a temperature of 37C for 5 minutes. After which,  we added the peptide pulsed APCs (10 APCs to 1 OT-1 T cell) and pelleted mixture of cells by centrifugation for 10 s at 460 rcf at room temperature.  The pellet was then incubated at a temperature of 37C for 10 minutes, followed by 15 minutes of chemical fixing with 2\% PFA on ice and then permeabilization with ice cold 90\% MeOH. Samples were kept at a temperature of -20C until labeling for FCM.

\subsubsection*{Experiment 2:} C57BL/6N splenocytes were pulsed for 2hr with 100nM SIINFEKL peptide, then irradiated (3000RAD), washed once and used as stimulator/feeder cells. OT-1 lymphocytes were harvested from auxiliary, brachial and inguinal lymph nodes as well as spleen (splenocytes were treated with ACK lysis buffer to remove red blood cells), and mixed with SIINFEKL-pulsed C57BL/6N splenocytes in complete RPMI. After two days, cells were expanded by diluting 2 fold into medium containing 100 pM IL-2. After four days, the cells were again expanded by 2 fold dilution into medium with IL-2. After one more day of culture, cells were harvested and spun through a 1.09 density Ficoll- Paque Plus gradient (GE Healthcare) to remove dead cells. Live cells were recovered, washed twice in complete medium and resuspended at 1 million/ml in complete medium with 100pM IL-2. Cells were used for experiments between 6 and 8 days after primary stimulation. The ppERK response to SRC and MEK inhibition was measured using primary OT-1 T-lymphocytes activated with RMA-S APCs. RMA-S cells were suspended in culture with 1 nM SIINFEKL peptide for 2 h at 37C, 5\% CO\textsubscript{2}, and on a rotator to guarantee mixing. During this time we labelled OT-1 cells with an amine-reactive dye, CTV, according to the manufacture’s protocol (Molecular Probes). We rested the OT-1 cells one hour after CTV staining, and then distributed them in a 96-well v-bottom plate. Each well was given various doses of SRC inhibitor and MEK inhibitor and kept at 37C for 5 min. Following the 5 min exposure to the inhibitors, we added the peptide pulsed RMA-S (10 RMA-S to 1 OT-1 T cell) and pelleted by centrifugation for 10 s at 460 rcf at room temperature. This step guaranteed that both cell types, OT-1 and RMA-S, came into contact. The cells were allowed to activate for 10 min, followed by fixing on ice in 2\% PFA, and then permeabilized and stored in 90\% MeOH at  -20C. APCs were pulsed with serial dilutions of OVA or variant peptides for 2 hr at 37C, then washed with T cell medium at the time of harvest, and resuspended with anti-CD8 (53-6.72)-Fab-coated T cells in their conditioned media in a V-bottom 96-well plate (Corning). Fab-coating was performed 10 min before cell use by incubating T cells with 10 $\mu$g/ml of Fab fragment. T:APC cell contacts were synchronized using a quick centrifugal spin (10s at 400g). Plates containing T:APC conjugates were placed on a water bath at 37C and incubated for 5 min. Supernatants were then discarded, T:APC conjugates disrupted by vortexing, and cells resuspended in ice-cold 4\% paraformaldehyde for 15 min. Cells were then permeabilized with ice-cold 90\% methanol for 15 min on ice, and washed twice with FACS buffer.

@@

\smallskip
In experiment 1, the T-cells were inhibited by the SRC inhibitor Dasatinib. In Fig.~\ref{fig:Experiment}(A) we see how a commercially-available analysis software (FlowJo \cite{FlowJo}) plots the distribution of ppERK in such an experiment, which clearly shows a bimodal structure. Fig.~\ref{fig:Experiment}(B) plots those same data, formally $P(\log I) := \hat{f}_{\log I}^{(8751)}$, the probability density estimated from the data by taking the logarithm of the intensity of 8751 cells and binning it in bins of size 0.2. Our procedure for estimating $P(\log I)$ gives the two modes as in FlowJo (red dots) whereas $Q(I)$ (derived using Eq.~\ref{eq:P2Q}) has a single mode (blue dots). It also agrees with the results of Hartigan's unimodality p-values, $p_U$, explained in more detail in the \emph{Supporting information}. We fit $P(\log I)$ as a Gaussian mixture. This is followed in Fig.~\ref{fig:Experiment}(C) by the same extrema analysis as in Fig.~\ref{fig:ExtremaSolutions}, revealing that indeed $Q(I)$ has a single maximum.
\smallskip

@@

\begin{figure}[!htb]
 \centering
%   \includegraphics[width=1\linewidth]{Fig7.tif}  
  \caption{{\bf Test for weak dependence of ERK1 and activation status.} Whereas independence implies that $P(\log ERK1) = P(\log ERK1 \vert Activation)$, in fact there is a weak dependence between them regardless of whether we employ a vertical (red, defined by threshold value $ppERK_*$) or a diagonal (grey) gate (the diagonal gate showing a weaker dependence); this is observed directly by noting that the different distributions in Fig.~\ref{S1_Fig} do not lie on top of each other. To quantify this difference, the inset shows the mutual information between $P(\log ERK1)$ and $P(\log ERK1 | Activation)$, with the circle pointing out the particular concentration of stimulus (out of all concentrations used in Ref.~\cite{Feinerman2008}) we chose to plot in this example. The chosen concentration has the highest mutual information, {\it i.e.}, the lowest ability to discern between the active and inactive states.}
  \label{S1_Fig}
\end{figure}

@@

%\subsection*{Relation between ERK1 level and activation status}
